### 2021-12-13 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@0fce0931a5ebc09dbe834c2b1cd1595c42b2a141 🚀

\hypertarget{illustration-on-empirical-food-webs}{%
\section{Illustration on empirical food
webs}\label{illustration-on-empirical-food-webs}}

In this section, we use a collection of XX empirical networks

\hypertarget{using-t-sbn-and-rdpg-to-simulate-arbitrarily-imperfect-classifier}{%
\subsection{Using t-SBN and RDPG to simulate arbitrarily imperfect
classifier}\label{using-t-sbn-and-rdpg-to-simulate-arbitrarily-imperfect-classifier}}

use t-SVD to only use some information

cumulative eigenvalues up to rank n give amount of information recovered

can be thresholded

how do values react to using fewer ranks

\hypertarget{effect-of-classifier-performance-on-metrics}{%
\subsection{Effect of classifier performance on
metrics}\label{effect-of-classifier-performance-on-metrics}}

\ldots{}

\hypertarget{numerical-experiments}{%
\section{Numerical experiments}\label{numerical-experiments}}

In the following section, we will generate random networks, and train
four binary classifiers (as well as an ensemble model using the sum of
the outputs) on 30\% of the interaction data. Networks are generated by
picking random generality \(g\) and vulnerability \(v\) traits for
\(S = 200\) species uniformly on the unit interval, and assigning an
interaction from species \(i\) to species \(j\) if
\(0.2g_i-\xi \le v_j \le 0.2g_i+\xi\), where \(\xi\) is a constant
regulating the connectance of the networks, and varies uniformly in
\([5\times 10^{-3}, 10^{-1}]\). This model gives fully interval networks
that are close analogues to the niche model (Williams \& Martinez,
2000), but has the benefit of only relying on two features
(\(g_i, v_j\)), and having the exact same rule for all interactions. It
is, therefore, a simple case which most classifiers should be able to
learn.

The training sample is composed of 30\% of the \(4\times 10^4\) possible
entries in the network, \emph{i.e.} \(n=12000\). Out of these
interactions, we pick a proportion \(\nu\) (the training set bias) to be
positive, so that the training set has \(\nu n\) interactions, and
\((1-\nu) n\) non-interactions. We vary \(\nu\) uniformly in \(]0,1[\).
This allows to evaluate how the measures of binary classification
performance respond to artificially rebalanced dataset for a given
network connectance. Note that both \(\xi\) and \(\nu\) are sampled from
a distribution rather than being picked on a grid; this is because there
is no direct relationship between the value of \(\xi\) and the
connectance of the simulated network, and therefore the precise value of
\(\xi\) is not relevant for the analysis of the results.

The dataset used for numerical experiments is composed of 64000 such
\((\xi, \nu)\) pairs, on which four learners are trained: a decision
tree regressor, a boosted regression tree, a ridge regressor, and a
random forest regressor. All models were taken from the \texttt{MLJ.jl}
package (Blaom et al., 2020; Blaom \& Vollmer, 2020) in Julia 1.7
(Bezanson et al., 2017). In order to pick the best adjacency matrix for
a given learner, we performed a thresholding approach using 500 steps on
predictions from the testing set, and picking the threshold that
maximized Youden's informedness, which is usually the optimized target
for imbalanced classification. During the thresholding step, we measured
the area under the receiving-operator characteristic (ROC-AUC) and
precision-recall (PR-AUC) curves, as measures of overall performance
over the range of returned values. We report the ROC-AUC and PR-AUC, as
well as a suite of other measures as introduced in the next section, for
the best threshold. The ensemble model was generated by summing the
predictions of all component models on the testing set (ranged in
\([0,1]\)), then put through the same thresholding process. The complete
code to run the simulations is given as an appendix; running the final
simulation required 4.8 core days (approx. 117 hours).

@@

@@

@@

\leavevmode\hypertarget{ref-Williams2000SimRul}{}%
Williams, R., \& Martinez, N. (2000). Simple rules yield complex food
webs. \emph{Nature}, \emph{404}, 180--183.
\href{http://userwww.sfsu.edu/\%20}{http://userwww.sfsu.edu/}

>>>>

\hypertarget{numerical-experiments-on-training-strategy}{%
\section{Numerical experiments on training
strategy}\label{numerical-experiments-on-training-strategy}}

In the following section, we will generate random bipartite networks
(this works without loss of generality on unipartite networks), and
train four binary classifiers (as well as an ensemble model using the
sum of ranged outputs from the component models) on 30\% of the
interaction data. Networks are generated by picking a random
infectiousness trait \(v_i\) for 100 species (from a \(B(6,8)\)
distribution), and a resistance trait \(h_j\) for 100 species (from a
\(B(2,8)\) distribution). There is an interaction between \(i\) and
\(j\) when \(v_i-\xi/2 \le h_j \le v_i+\xi/2\), where \(\xi\) is a
constant regulating the connectance of the network (there is an almost
1:1 relationship between \(\xi\) and connectance), and varies uniformly
in \([0.05, 0.35]\). This model gives fully interval networks that are
close analogues to the bacteria--phage model of Weitz et al. (2005),
with both a modular structure and a non-uniform degree distribution.
This model is easy to learn: when trained with features
\([v_i, h_j, \text{abs}(v_i, h_j)] ^T\) to predict the interactions
between \(i\) and \(j\), all four models presented below were able to
reach almost perfect predictions all the time (data not presented here)
-- this is in part because the rule is fixed for all interactions. In
order to make the problem more difficult to solve, we use \([v_i, h_j]\)
as a feature vector, and therefore the models will have to uncover that
the rule for interaction is \(\text{abs}(v_i, h_j) \le \xi\).

The training sample is composed of 30\% of the \(10^4\) possible entries
in the network, \emph{i.e.} \(n=3000\). Out of these interactions, we
pick a proportion \(\nu\) (the training set bias) to be positive, so
that the training set has \(\nu n\) interactions, and \((1-\nu) n\)
non-interactions. We vary \(\nu\) uniformly in \(]0,1[\). This allows to
evaluate how the measures of binary classification performance respond
to artificially rebalanced dataset for a given network connectance. The
rest of the dataset (\(n=7000\) pairs of species) is used as a testing
set, on which all furher measures are calculated. Note that although the
training set is balanced, the testing set is not, and retains (part of)
the imbalance of the original data.

The dataset used for numerical experiments is composed of 64000 such
\((\xi, \nu)\) pairs, on which four machines are trained: a decision
tree regressor, a boosted regression tree, a ridge regressor, and a
random forest regressor. All models were taken from the \texttt{MLJ.jl}
package (Blaom et al., 2020; Blaom \& Vollmer, 2020) in Julia 1.7
(Bezanson et al., 2017). All machines use the default parameterization;
this is an obvious deviation from best practices, as the hyperparameters
of any machine require training before its application on a real
dataset. As we use 64000 such datasets, this would require 256000 unique
instances of tweaking thehyperparameters, which is not realistic.
Therefore, we assume that the default parameterizations are comparable
across networks. All machines return a quantitative prediction, usually
(but not necessarilly) in \([0,1]\), which is proportional (but not
necessarilly linearly) to the probability of an interaction between
\(i\) and \(j\).

In order to pick the best adjacency matrix for a given trained machine,
we performed a thresholding approach using 500 steps on predictions from
the testing set, and picking the threshold that maximized Youden's
informedness, which is usually the optimized target for imbalanced
classification. During the thresholding step, we measured the area under
the receiving-operator characteristic (ROC-AUC) and precision-recall
(PR-AUC) curves, as measures of overall performance over the range of
returned values. We report the ROC-AUC and PR-AUC, as well as a suite of
other measures as introduced in the next section, for the best
threshold. The ensemble model was generated by summing the predictions
of all component models on the testing set (ranged in \([0,1]\)), then
put through the same thresholding process. The complete code to run the
simulations is given as an appendix; running the final simulation
required 4.8 core days (approx. 117 hours).

@@

\hypertarget{do-better-classification-accuracy-result-in-more-realistic-networks}{%
\section{Do better classification accuracy result in more realistic
networks?}\label{do-better-classification-accuracy-result-in-more-realistic-networks}}

@@

\leavevmode\hypertarget{ref-Weitz2005CoeArm}{}%
Weitz, J. S., Hartman, H., \& Levin, S. A. (2005). Coevolutionary arms
races between bacteria and bacteriophage. \emph{Proceedings of the
National Academy of Sciences of the United States of America},
\emph{102}(27), 9535--9540.
\url{https://doi.org/10.1073/pnas.0504062102}

@@
